---
title: 'Introduction to clustifyr'
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyr-overview}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r "knitr options", echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  fig.align = "center",
  comment = "#>"
)
```

## Why use `clustifyr`?

Single cell transcriptomes are difficult to annotate without extensive knowledge of the underlying biology of the system in question. Even with this knowledge, accurate identification can be challenging due to the lack of detectable expression of common marker genes defined by bulk RNA-seq, flow cytometry, other single cell RNA-seq platforms, etc.

`clustifyr` solves this problem by providing functions to automatically annotate single cells or clusters using bulk RNA-seq data or marker gene lists (ranked or unranked). Additional functions allow for exploratory analysis of calculated similarities between single cell RNA-seq datasets and reference data.

## Installation

To install `clustifyr` BiocManager must be installed.

```{"Installation"}
install.packages("BiocManager")

BiocManager::install("clustifyr")
```

## A simple example: 10x Genomics PBMCs

In this example, we take a 10x Genomics 3' scRNA-seq dataset from peripheral blood mononuclear cells (PBMCs) and annotate the cell clusters (identified using `Seurat`) using scRNA-seq cell clusters assigned from a CITE-seq experiment. The cell types in this dataset have been previously annotated, so this is a good dataset to use for testing the accuracy of `clustifyr`.

```{r "Load data"}
library(clustifyr)
library(ggplot2)
library(cowplot)

# Matrix of normalized single-cell RNA-seq counts
full_pbmc_matrix <- clustifyrdata::pbmc_matrix

# meta.data table containing cluster assignments for each cell
# The table that we are using also contains the known cell identities in the "classified" column
full_pbmc_meta <- clustifyrdata::pbmc_meta
```

## Calculate correlation coefficients

To identify cell types, the `clustifyr()` function requires several inputs:

* `input`: a Seurat object or a matrix of normalized single-cell RNA-seq counts
* `metadata`: a meta.data table containing the cluster assignments for each cell (not required if a Seurat object is given)
* `ref_mat`: a reference matrix containing RNA-seq expression data for each cell type of interest
* `query_genes`: a list of genes to use for comparison (optional)

When using a matrix of scRNA-seq counts `clustifyr()` will return a matrix of correlation coefficients for each cell type and cluster, with the rownames corresponding to the cluster number.

```{r "Run clustifyr()"}
# Calculate correlation coefficients for each cluster (spearman by default)
vargenes <- pbmc_vargenes[1:500]

res <- clustify(
  input = full_pbmc_matrix,        # matrix of normalized scRNA-seq counts (or a Seurat object)
  metadata = full_pbmc_meta,       # meta.data table containing cell clusters
  ref_mat = cbmc_ref,              # matrix of RNA-seq expression data for each cell type
  query_genes = vargenes,          # list of highly varible genes identified with Seurat
  cluster_col = "seurat_clusters"  # name of column in meta.data containing cell clusters
)

# Peek at correlation matrix
res[1:5, 1:5]
```

To visualize the `clustifyr()` results we can use the `plot_cor_heatmap()` function to plot the correlation coefficients for each cluster and each cell type.

```{r "Create correlation heatmap", fig.height = 5, fig.width = 7}
# Create heatmap of correlation coefficients using clustifyr() output
plot_cor_heatmap(cor_mat = res)
```

## Plot cluster identities and correlation coefficients

`clustifyr` also provides functions to overlay correlation coefficients on pre-calculated tSNE embeddings (or those from any other dimensionality reduction method). 

```{r "Overlay corr coefficients on UMAP", fig.height = 3.5, fig.width = 9}
# Overlay correlation coefficients on UMAPs for the first two cell types
corr_umaps <- plot_cor(
  cor_mat = res,                      # matrix of correlation coefficients from clustifyr()
  metadata = full_pbmc_meta,          # meta.data table containing UMAP or tSNE data
  data_to_plot = colnames(res)[1:2],  # name of cell type(s) to plot correlation coefficients
  cluster_col = "seurat_clusters"     # name of column in meta.data containing cell clusters
)

plot_grid(
  plotlist = corr_umaps,
  rel_widths = c(0.47, 0.53)
)
```

<br>

The `plot_best_call()` function can be used to label each cluster with the cell type that gives the highest corelation coefficient. Using the `plot_tsne()` function, we can also plot the known identities of each cluster, which were stored in the "classified" column of the meta.data table. The plots below show that the highest correlations between the reference RNA-seq data and the 10x Genomics scRNA-seq dataset are restricted to the correct cell clusters.

```{r "Label clusters", fig.height = 5.5, fig.width = 12}
# Label clusters with clustifyr cell identities
clustifyr_types <- plot_best_call(
  cor_mat = res,              # matrix of correlation coefficients from clustifyr()
  metadata = full_pbmc_meta,  # meta.data table containing UMAP or tSNE data
  do_label = TRUE,            # should the feature label be shown on each cluster?
  do_legend = FALSE,          # should the legend be shown?
  cluster_col = "seurat_clusters"
) +
  ggtitle("clustifyr cell types")

# Compare clustifyr results with known cell identities
known_types <- plot_tsne(
  data = full_pbmc_meta,   # meta.data table containing UMAP or tSNE data
  feature = "classified",  # name of column in meta.data to color clusters by
  do_label = TRUE,         # should the feature label be shown on each cluster?
  do_legend = FALSE        # should the legend be shown?
) +
  ggtitle("Known cell types")

plot_grid(known_types, clustifyr_types)
```

## Classify cells using known marker genes

The `clustify_lists()` function allows cell types to be assigned based on known marker genes. This function requires a table containing markers for each cell type of interest. Cell types can be assigned using several statistical tests including, hypergeometric, Jaccard, Spearman, and GSEA.

```{r "clustifyr with gene lists", fig.height = 4, fig.width = 6}
# Take a peek at marker gene table
cbmc_m

# Available metrics include: "hyper", "jaccard", "spearman", "gsea"
list_res <- clustify_lists(
  input = full_pbmc_matrix,         # matrix of normalized single-cell RNA-seq counts
  metadata = full_pbmc_meta,        # meta.data table containing cell clusters
  cluster_col = "seurat_clusters",  # name of column in meta.data containing cell clusters
  marker = cbmc_m,                  # list of known marker genes
  metric = "hyper"                  # test to use for assigning cell types
)

# view as heatmap, or plot_best_call
plot_cor_heatmap(
  cor_mat = list_res,            # matrix of correlation coefficients from clustify_lists()
  col = not_pretty_palette,      # colors to use for heatmap
  cluster_rows = FALSE,          # cluster by row?
  cluster_columns = FALSE,       # cluster by column?
  legend_title = "-log10(padj)"  # title of heatmap legend
)
```

## Direct handling of `seurat` v2 and v3 objects

`clustifyr()` can also use a `Seurat` object as input and return a new `Seurat` object with the cell types added as a column in the meta.data. 

```{r}
res <- clustify(
  input = s_small,        # a Seurat object
  ref_mat = cbmc_ref,     # matrix of RNA-seq expression data for each cell type
  cluster_col = "res.1"   # name of column in meta.data containing cell clusters
)
```

## Session info

```{r}
sessionInfo()
```

