---
title: "Compare Methods"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyR-compare-methods}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

This vignette illustrates how one can explore how well different feature selection methods work (which gene selection methods yields the most useful correlations). We have 4 different feature selection methods 

1) Highly variable genes via Seurat
2) Markers of each cluster
3) M3Drop identified genes 
4) PCA based method

## Load required example data sets

```{r setup, echo=T}
library(clustifyR)

#Assign pre-labelled names to single cell clusters
names(pbmc4k_avg) <- c(
   "CD4 T cells, 1",
   "CD14+ Monocytes",
   "B cells",
   "CD4 T cells, 2",
   "CD8 T cells",
   "NK cells",
   "Dendritic cells, 1",
   "FCGR3A+ Monocytes",
   "Dendritic Cells, 2?",
   "Megakaryocytes"
)
```

## Function to create correlation matrix 

* Row names are the single cell RNA pre-assigned tissue labels.
* Column names are the bulk RNA sample names.
* Values are the pearson correlations between the average single cell RNA expression and the bulk RNA sample expression, for a given set of genes (determined by the feature selection methods).

```{r function_corr_matrix}

cor_mat <- clustify(expr_mat = pbmc4k_matrix,
               metadata = pbmc4k_meta,
               bulk_mat = pbmc_bulk_matrix,
               query_gene_list = pbmc4k_vargenes,
               cluster_col = "classified",
               compute_method = corr_coef)

knitr::kable(round(cor_mat, 2))

```

## Function to create best match matrix 

This function takes a correlation matrix, and determines which bulk RNA sample is the best match (1) for a single cell RNA tissue

```{r function_best_match, echo=T}
get_best_match_matrix <- function(correlation_matrix) {
  
  best_mat <- as.data.frame(t(apply(correlation_matrix, 1, function(x) x - max(x))))
  best_mat[best_mat==0]="1"
  best_mat[best_mat!="1"]="0"

  return (best_mat)
}

best_match_matrix <- get_best_match_matrix(cor_mat)

knitr::kable(best_match_matrix)
```

## Summary of best matches

Using the above functions, we can now summarize and compare and contrast the best matches for each of the different feature selection methods.

```{r summary_table, echo=T}
vargenes_cor_mat <- clustify(expr_mat = pbmc4k_matrix,
               metadata = pbmc4k_meta,
               bulk_mat = pbmc_bulk_matrix,
               query_gene_list = pbmc4k_vargenes,
               cluster_col = "classified",
               compute_method = corr_coef)

vargenes_best_mat <- get_best_match_matrix(vargenes_cor_mat)

markers_cor_mat <- clustify(expr_mat = pbmc4k_matrix,
               metadata = pbmc4k_meta,
               bulk_mat = pbmc_bulk_matrix,
               query_gene_list = pbmc4k_markers$gene,
               cluster_col = "classified",
               compute_method = corr_coef)

markers_best_mat <- get_best_match_matrix(markers_cor_mat)


m3drop_cor_mat <- clustify(expr_mat = pbmc4k_matrix,
               metadata = pbmc4k_meta,
               bulk_mat = pbmc_bulk_matrix,
               query_gene_list = pbmc4k_markers_M3Drop$Gene,
               cluster_col = "classified",
               compute_method = corr_coef)

m3drop_best_mat <- get_best_match_matrix(m3drop_cor_mat)


pca_cor_mat <- clustify(expr_mat = pbmc4k_matrix,
               metadata = pbmc4k_meta,
               bulk_mat = pbmc_bulk_matrix,
               query_gene_list = pbmc_pca,
               cluster_col = "classified",
               compute_method = corr_coef)

pca_best_mat <- get_best_match_matrix(pca_cor_mat)

get_best_str <- function(name, best_mat, cor_mat) {
  if (sum(as.numeric(best_mat[,name])) > 0) {
      best.names <- rownames(best_mat)[which(best_mat[,name]==1)]  
      best.cor <- round(cor_mat[which(best_mat[,name]==1),name],2)
      for (i in 1:length(best.cor)) {
        if (i == 1) {
          str <- paste0(best.names[i], " (", best.cor[i], ") ") 
        } else {
            str <- paste0(str, "; ", best.names[i], " (", best.cor[i], ") ") 
        }
      }
  } else {
    str <- ""
  }
  return(str)
}

summary <- data.frame()
bulk_rna_sample_names <- colnames(pbmc_bulk_matrix)
for (name in bulk_rna_sample_names) {
  variable_genes <- get_best_str(name, vargenes_best_mat, vargenes_cor_mat)
  marker_genes <- get_best_str(name, markers_best_mat, markers_cor_mat)
  m3drop_genes <- get_best_str(name, m3drop_best_mat, m3drop_cor_mat)
  pca_genes <- get_best_str(name, pca_best_mat, pca_cor_mat)
  if (paste0(variable_genes, marker_genes, m3drop_genes, pca_genes) != "") {
      row <- data.frame(
        variable_genes,
        marker_genes,
        m3drop_genes,
        pca_genes
      )
      rownames(row) <- name
      summary <- rbind(summary,row)    
  }
}

knitr::kable(summary)

```



