---
title: "Negative control analysis"
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyR-negative-controls}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(clustifyR)
library(ComplexHeatmap)
library(circlize)
library(recount)
library(dplyr)
```

## Download data from recount

We provide a utility function to download data from the recount2 database into an easy to process format. Here we will download a study that contains a cell line to examine as a negative control. 

```{r}
source(system.file("dl_recount.R", 
                   package = "clustifyR"))

cell_line_data <- dl_recount("SRP055105")

cell_line_data$read_counts[1:5, 1:3]
cell_line_data$meta_data[1:3, ]
unique(cell_line_data$meta_data$transfection)
```

Next we'll add in some informative column names for the control samples

```{r}
good_libs <- cell_line_data$meta_data %>% 
  filter(str_detect(transfection, "Control"))

good_libs$id <- paste0(good_libs$`cell line`, 
                       " rep ", 1:length(good_libs$`cell line`))
  
cell_line_counts <- cell_line_data$read_count[, good_libs$run] 
colnames(cell_line_counts) <- good_libs$id

cell_line_counts[1:5, 1:3]
```

Now we'll add in the 293T data into the PBMC experiment and recompute correlation metrics. 

```{r}
shared_genes <- intersect(rownames(pbmc_bulk_matrix),
                          rownames(cell_line_counts))

new_bulk_matrix <- cbind(pbmc_bulk_matrix[shared_genes, ],
                         cell_line_counts[shared_genes, ])

new_bulk_matrix[1:5, c(1, 20)]
```


```{r, fig.height=8, fig.width=12}
# run correlation (pearson by default)
res <- run_cor(pbmc4k_matrix,
        pbmc4k_meta,
        new_bulk_matrix,
        pbmc4k_vargenes,
        if_permute = F,
        metadata = "classified",
        compute_method = corr_coef)

# view heatmap of correlation coefficients
Heatmap(res, colorRamp2(c(0, 1), c("white", "black")))
```
