---
title: 'Support for various single cell object formats'
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyr-formats}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r "knitr options", echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

`clustifyr` is aimed for general use and flexibility in various analysis pipelines. While its core code functions with matrices and dataframes, and hence has minimal package dependencies, wrapper functions are provided for further convenience of the user.

## Direct handling of `seurat` objects, v2 and v3, extracting the required data, and inserting called cell types directly into the objects

```{r}
library(clustifyr)

# can even returning a new object with `type` added to metadata.
res <- clustify(
  input = s_small,
  cluster_col = "res.1",
  ref_mat = cbmc_ref,
  seurat_out = TRUE
)

# or return a correlation matrix, if seurat_out is set to F
res2 <- clustify_lists(
  input = s_small,
  marker = pbmc_markers,
  marker_inmatrix = FALSE,
  cluster_col = "res.1",
  seurat_out = FALSE
)
```

## Other scRNA-seq object formats are easily supported, by adding defaults to a lookup file `object_loc_lookup`, currently including `SCE`, `URD`, `CDS`, `FSCE`.

```{r}
# location of data in object in this format
object_loc_lookup

# adding seurat3 locations
object_loc_lookup$Seurat <- c(
  expr = "input@assays$RNA@data",
  meta = "input@meta.data",
  var = "input@assays$RNA@var.features",
  col = "RNA_snn_res.1"
)
```

For example, sample code for `scrunchy` is as easy as:

```{"example"}
res <- clustify_nudge(
  input = fsce_small,
  ref_mat = cbmc_ref,
  marker = cbmc_m,
  query_genes = pbmc_vargenes,
  cluster_col = "k_cluster"
)
```

## Making new references from objects
Making scRNA-seq objects into references is assisted by wrappers as well.

```{r, messag = T}
# for both v2 and v3
seurat_ref <- seurat_ref(
  seurat_object = s_small3,
  cluster_col = "RNA_snn_res.1"
)

# and other object types with lookup
object_ref <- object_ref(input = s_small3)
```
