---
title: 'Support for various single cell object formats'
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyR-formats}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

`clustifyR` is aimed for general use and flexibility in various analysis pipelines. While its core code functions with matrices and dataframes, and hence has minimal package dependencies, wrapper functions are provided for further convenience of the user.

## Direct handling of `seurat` objects, v2 and v3, extracting the required data, and inserting called cell types directly into the objects

```{r, eval = F, warning = F, message = F}
library(clustifyR)
# can even returning a new object with `type` added to metadata.
res <- clustify(
  input = s_small,
  cluster_col = "res.1",
  ref_mat = pbmc_bulk_matrix,
  seurat_out = T
)

# or return a correlation matrix, if seurat_out is set to F
res2 <- clustify_lists(
  s_small,
  per_cell = F,
  marker = pbmc4k_markers,
  marker_inmatrix = F,
  cluster_col = "res.1",
  seurat_out = F
)

# and combined ref matrix and marker info
res3 <- clustify_nudge(
  input = s_small3,
  ref_mat = cbmc_ref,
  marker = cbmc_m,
  cluster_col = "RNA_snn_res.1",
  seurat_out = F
)
```

## Other scRNA-seq object formats are easily supported, by adding defaults to a lookup file `object_loc_lookup`

```{r lookup, warning = F, message = F}
# location of data in object in this format
library(clustifyR)
object_loc_lookup

# adding seurat3 locations
object_loc_lookup$Seurat <- c(expr = "input@assays$RNA@data", 
                              meta = "input@meta.data", 
                              var = "input@assays$RNA@var.features",
                              col = "RNA_snn_res.1"
                              )
```

For example, sample code for `scrunchy` is as easy as:
```{r scrunchy, warning = F, message = F, eval = F}
res <- clustify_nudge(
  fsce_small,
  cbmc_ref,
  cbmc_m,
  query_genes = pbmc4k_vargenes,
  cluster_col = "k_cluster"
)
```
