---
title: 'Visualization of clustifyr results'
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{clustifyR-viz}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

## Plotting tSNE and UMAP results

`clustifyr` provides several functions to plot tSNE or UMAP results. The `plot_tsne()` function will plot tSNE or UMAP data using a meta.data table and can color the cells based on cluster identity.

```{r "Create tSNE", fig.height = 8, fig.width = 10}
library(clustifyR)

# Matrix of normalized single-cell RNA-seq counts
pbmc4k_matrix

# meta.data table containing cluster assignments for each cell 
pbmc4k_meta[1:5, ]

# Create tSNE and color cells based on cluster ID
plot_tsne(
  x = "tSNE_1",             # name of column in the meta.data containing the data to plot on x-axis
  y = "tSNE_2",             # name of column in the meta.data containing the data to plot on y-axis
  data = pbmc4k_meta,       # meta.data table containing cluster assignments for each cell
  feature = "cluster"       # name of column in meta.data to color cells by
)
```

Cells can also be colored based on the expression level of a gene or list of genes using the `plot_gene()` function.   

```{r "Overlay gene expression"}
# Create tSNE and color cells based on gene expression
plot_gene(
  x = "tSNE_1",                # name of column in the meta.data containing the data to plot on x-axis
  y = "tSNE_2",                # name of column in the meta.data containing the data to plot on y-axis
  expr_mat = pbmc4k_matrix,    # matrix of normalized single-cell RNA-seq counts
  metadata = pbmc4k_meta,      # meta.data table containing cluster assignments for each cell
  genes = c("CD79B", "CD8A"),  # vector of gene names to color cells
  cell_col = "rn"              # name of column in meta.data containing the cell IDs
)
```

## Visualizing `clustifyr()` results

The `clustifyr()` function outputs a matrix of correlation coefficients and `clustify_lists()` and `clustify_nudge()` output positive scores. `clustifyr` provides built-in functions to help visualize these results.

Cell type assignments can be assessed by plotting the `clustifyr()` correlation matrix as a heatmap using the `plot_cor_heatmap()` function.

```{r "Correlation heatmap"}
# Run clustifyr()
res <- clustify(
  input = pbmc4k_matrix,          # matrix of normalized single-cell RNA-seq counts
  metadata = pbmc4k_meta,         # meta.data table containing cluster assignments for each cell
  ref_mat = cbmc_ref,             # reference matrix containing bulk RNA-seq data for each cell type
  query_genes = pbmc4k_vargenes,  # list of highly varible genes identified with Seurat
  cluster_col = "cluster"         # name of column in meta.data containing cell clusters
)

# Create heatmap using the clustifyr correlation matrix
plot_cor_heatmap(
  cor_matrix = res  # matrix of correlation coefficients from clustifyr()
)
```

The `plot_cor()` function can also be used to create a tSNE for each cell type of interest and color the cells based on the correlation coefficients.

```{r "Overlay corr coefficients", fig.height = 5.5, fig.width = 12}
# Create a tSNE for each cell type of interest and color cells based on correlation coefficients
plot_cor(
  x = "tSNE_1",                       # name of column in the meta.data containing the data to plot on x-axis
  y = "tSNE_2",                       # name of column in the meta.data containing the data to plot on y-axis
  cor_matrix = res,                   # matrix of correlation coefficients from clustifyr()
  metadata = pbmc4k_meta,             # meta.data table containing cluster assignments for each cell
  data_to_plot = colnames(res)[1:2],  # name of cell type(s) to plot correlation coefficients
  cluster_col = "cluster"             # name of column in meta.data containing cell clusters
)
```

Cell clusters can also be labeled using the `plot_best_call()` function, which takes the correlation matrix and labels cell clusters with the cell type that has the highest correlation coefficient.

```{r "Label clusters with cell types", fig.height = 8, fig.width = 10}
# Create tSNE and label clusters with the cell type that has the highest correlation coefficient
plot_best_call(
  cor_matrix = res,        # matrix of correlation coefficients from clustifyr()
  metadata = pbmc4k_meta,  # meta.data table containing UMAP or tSNE data
  do_label = TRUE,         # should the feature label be shown on each cluster?
  do_legend = FALSE        # should the legend be shown?
)
```

## Assessing `clustifyr()` accuracy

The `clustifyr()` results can also be evaluated by over-clustering the data and comparing the cell type assignments before and after over-clustering. This is accomplished using the `overcluster_test()` function. The cell type assignments should be similar with and without over-clustering.

```{r "Overclustering test", fig.height = 8, fig.width = 10}
# Overcluster cells and compare cell type assignments with and without over-clustering
overcluster_test(
  expr = pbmc4k_matrix,     # matrix of normalized single-cell RNA-seq counts
  metadata = pbmc4k_meta,   # meta.data table containing UMAP or tSNE data
  ref_mat = cbmc_ref,       # reference matrix containing bulk RNA-seq data for each cell type      
  cluster_col = "cluster",  # name of column in meta.data containing cell clusters
  n = 5                     # expand cluster number n-fold for overclustering
)
```

The cell types from the bulk RNA-seq reference matrix can also be mixed together using the `make_comb_ref()` function to assess the specificity of the cell type assignments. If a cluster shows a higher correlation when using the mixed reference matrix, this suggests that the cluster contains multiple cell types.

```{r "Create combined reference", fig.height = 8, fig.width = 10}
# Create reference containing different combinations of the bulk RNA-seq data
comb_ref <- make_comb_ref(
  ref_mat = cbmc_ref  # reference matrix containing bulk RNA-seq data for each cell type      
)

# Peek at the new combined reference
comb_ref[1:5, 1:5]

# Run clustifyr() using the combined reference
comb_res <- clustify(
  input = pbmc4k_matrix,          # matrix of normalized single-cell RNA-seq counts
  metadata = pbmc4k_meta,         # meta.data table containing cluster assignments for each cell
  ref_mat = comb_ref,             # reference matrix containing bulk RNA-seq data for each cell type
  query_genes = pbmc4k_vargenes,  # list of highly varible genes identified with Seurat
  cluster_col = "cluster"         # name of column in meta.data containing cell clusters
)

# Create tSNE and label clusters with the assigned cell types from the combined reference
plot_best_call(
  cor_matrix = comb_res,   # matrix of correlation coefficients from clustifyr()
  metadata = pbmc4k_meta,  # meta.data table containing UMAP or tSNE data
  do_label = TRUE,         # should the feature label be shown on each cluster?
  do_legend = FALSE        # should the legend be shown?
)
```

## Plotting other attributes from the meta.data table 

Visualization of other attributes shared in the metadata between ref and query by `plot_cols`, such as nGene, nUMI, mt_percentage, as another way of identity confirmation after `clustify`. Certain cell types have distinct patterns, more genes detected, for example.

```{r "Plot other attributes", fig.height = 8, fig.width = 10}
# plot_cols(
#   metadata = pbmc_meta,
#   cluster_col = "seurat_clusters",
#   cluster_col_called = "classified",
#   metadata_ref = "...",
#   cluster_col_ref = "...",
#   plot_col_ref = "..."
# )
```









```{r "GSEA top hits", eval = F, include = F}
# ARE WE KEEPING THIS??

# gene lists can be converted from gmt files (Broad)
list_reactome <- gmt_to_list(here("c2.cp.reactome.v6.2.symbols.gmt"))

full_pbmc4k_avg <- average_clusters(
  full_pbmc4k_matrix, 
  full_pbmc4k_meta, 
  cluster_col = "classified"
)

res <- plot_pathway_gsea(
  full_pbmc4k_avg[,1:2], 
  list_reactome, 
  topn = 3
)

# returns dataframe of gsea analysis
res[[1]][1:2,1:2]
# and plot
res[[2]]
```

```{r, eval = F, include = F}
# plot_call(
#   cor_matrix = res, 
#   metadata = full_pbmc4k_meta,
#   data_to_plot = colnames(res)[1:2],
#   x = "tSNE_1",
#   y = "tSNE_2"
#   
# )

plot_cols(
  metadata = full_pbmc4k_meta,
  cluster_col = "cluster",
  cluster_col_called = ""
)
```






